---
title: "Exploring Air Quality"
author: "Daniela"
date: "5/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Data

The data came from the Air Quality Historical Data Platform, an open data framework which gathers data from worldwide EPA monitering programs.

```{r cars}
#airqualitybeijing.csv <- #"/home/CAMPUS/mwl04747/github/China_air_pollution_COVID/beijing-air-quality.csv"
airqualitybeijing.csv = "/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/beijing-air-quality.csv"

AQbeijing <- read.csv(airqualitybeijing.csv)

str(AQbeijing)

```

## Scatter Plot

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(AQbeijing$pm10)
```

Dates aren't correct -- Index...

## Next Steps

- Fix Dates

 - as.Date()
 
```{r fixdates}
# Fix the dates!
AQbeijing$NewDate = as.Date(AQbeijing$date)
str(AQbeijing)
head(AQbeijing)
```


```{r}
plot(pm25 ~ NewDate, data=AQbeijing)

```


 
- Select Data Before and After Lockdown
- Different for each city? Yes 
- 80 days

  - subset()
  
  
```{r subset}
# Dates for Beijing: "2020-02-10" Start

AQbeijingLockDown = as.Date("2020-02-10"); str(AQbeijingLockDown)

AQbeijingBefore = subset(AQbeijing, 
        subset=(NewDate > AQbeijingLockDown-80 &
                NewDate < AQbeijingLockDown))
AQbeijingAfter = subset(AQbeijing, 
        subset=(NewDate >= AQbeijingLockDown &
                NewDate < AQbeijingLockDown + 80))

plot(pm25~NewDate, AQbeijingBefore)
plot(pm25~NewDate, AQbeijingAfter)

```

- t-test Before and After
  - t.test()
  
  
```{r ttest}
#run ttest
t.test(AQbeijingBefore$pm25, AQbeijingAfter$pm25)


```

```{r ttest}
#extract mean_before, mean_after, p-value
ttest = t.test(AQbeijingBefore$pm25, AQbeijingAfter$pm25)
ttest$estimate
ttest$p.value


#export whole ttest
capture.output(ttest, file = "Welch_output.csv", append = TRUE)

#export just city, estimate, pvalue
estimate = ttest$estimate
pvalue = ttest$p.value
city = "Beijing"
capture.output(city, estimate, pvalue, file = "output.csv", append = TRUE)



#dump(city, estimate, pvalue, file = "output.csv", append=TRUE)

#save(city, estimate, pvalue, "output.csv")

#write.csv("Beijing", file="output.csv", row.names=FALSE, append=TRUE)


#write.csv(estimate, pvalue, file = "output.csv")

#save(estimate, pvalue, list = character(),
     #file = ("output.csv")

#save(estimate, pvalue, file = "output.RData")
#head("output.RData")
```

#repeating import data, scatterplot, fix dates, welch 2-sample t test, output for other cities.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Data

The data came from the Air Quality Historical Data Platform, an open data framework which gathers data from worldwide EPA monitering programs.

```{r cars}
#airqualitybeijing.csv <- #"/home/CAMPUS/mwl04747/github/China_air_pollution_COVID/beijing-air-quality.csv"


xianning.csv = "/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/data/e/xianning-air-quality.csv"

AQ <- read.csv(xianning.csv)

str(AQ)

```

## Scatter Plot

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(AQ$pm25)
```

Dates aren't correct -- Index...

## Next Steps

- Fix Dates

 - as.Date()
 
```{r fixdates}
# Fix the dates
AQ$NewDate = as.Date(AQ$date)
str(AQ)
head(AQ)
```


```{r}
plot(pm25 ~ NewDate, data=AQ)

```


 
- Select Data Before and After Lockdown
- Different for each city? Yes 
- 80 days

  - subset()
  
  
```{r subset}
# Dates for city: "yyyy-mm-dd" Start Lockdown

AQLockDown = as.Date("2020-01-24"); str(AQLockDown)

AQxianningBefore = subset(AQ, 
        subset=(NewDate > AQLockDown-80 &
                NewDate < AQLockDown))
AQxianningAfter = subset(AQ, 
        subset=(NewDate >= AQLockDown &
                NewDate < AQLockDown + 80))

plot(pm25~NewDate, AQxianningBefore) 
plot(pm25~NewDate, AQxianningAfter)

```

- t-test Before and After
  - t.test()
  
  
```{r ttest}
#run ttest
t.test(AQxianningBefore$pm25, AQxianningAfter$pm25)


```

```{r ttest}
#extract mean_before, mean_after, p-value
ttest = t.test(AQxianningBefore$pm25, AQxianningAfter$pm25)
ttest$estimate
ttest$p.value


#export whole ttest
capture.output(ttest, file = "Welch_output.csv", append = TRUE)

#export just city, estimate, pvalue
estimate = ttest$estimate
pvalue = ttest$p.value
city = "Xianning"
capture.output(city, estimate, pvalue, file = "output.csv", append = TRUE)



#dump(city, estimate, pvalue, file = "output.csv", append=TRUE)

#save(city, estimate, pvalue, "output.csv")

#write.csv("Beijing", file="output.csv", row.names=FALSE, append=TRUE)


#write.csv(estimate, pvalue, file = "output.csv")

#save(estimate, pvalue, list = character(),
     #file = ("output.csv")

#save(estimate, pvalue, file = "output.RData")
#head("output.RData")
```








- BACI "Before and Ater Control-Impact"
  - aov() 
  
  
visualization:

- time series -- with abreak B/A
- box plot B/A



```{r}
# cityname, date, pm25: two files: one before, one after lockdown starts
#excluding two problematic files: tianjin and xianning

city_size_lockdowndate.csv = "/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/city_size_lockdowndate.csv"

lockdowndate <- read.csv(city_size_lockdowndate.csv)

library(dplyr)
library(readr)
files <- list.files(path = "/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/data", pattern = "*.csv", full.names = T)

tbl.csv <- sapply(files, read_csv, simplify=FALSE) %>% 
bind_rows(.id = "id")

#write tbl as csv
write.csv(tbl, "tbl.csv")
head(tbl.csv)

#trying to include xianning and tianjin, which have no O3 values.
files2 <- list.files(path = "/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/data/tianjinandxianning/", pattern = "*.csv", full.names = T)
tbl_test <- sapply(files2, read_csv, simplify=FALSE) %>% 
  # use colClasses=c("o3"="numeric") with function read.csv
bind_rows(.id = "id")
write.csv(tbl_test, "tblnew.csv")
head(tbl_test.csv)

#rename cities in tbl from id
#renamed using command line: sed 's/"\/home\/CAMPUS\/dapa2019\/R\/China_air_pollution_COVID\/data\/baotou-air-quality\.csv"/baotou/g' tbl.csv>tbl2.csv
#renamed cities are in tbl_28cities.csv and tbl_2cities.csv. The 2cities file includes only tianjin and xianning.

#merge csv files: 28cities and 2cities

##files_all <- list.files(path = "/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/tbl/", pattern = "*cities.csv", full.names = T)

#tbl_all <- sapply(files_all, read_csv, simplify=FALSE) %>% 
#bind_rows(.id = "id")
#write.csv(tbl_all, "tbl_all")
#head(tbl_all.csv)


#all_data <- rbind.fill(files_all)


tbl_big <-read.csv("/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/tbl/tbl_28cities.csv",head=T)
tbl_small <-read.csv("/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/tbl/tbl_2cities.csv",head=T)
tbl_all = merge(tbl_big,tbl_small)
head(tbl_all) #gives 0 rows


#merge lockdown dates with imported data


#window to cut dates out

#create two big files


# ttest for before vs after




```


