---
title: "Exploring Air Quality"
author: "Daniela"
date: "5/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Data

The data came from the Air Quality Historical Data Platform, an open data framework which gathers data from worldwide EPA monitering programs.

```{r cars}
#airqualitybeijing.csv <- #"/home/CAMPUS/mwl04747/github/China_air_pollution_COVID/beijing-air-quality.csv"
airqualitybeijing.csv = "/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/beijing-air-quality.csv"

AQbeijing <- read.csv(airqualitybeijing.csv)

str(AQbeijing)

```

## Scatter Plot

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(AQbeijing$pm10)
```

Dates aren't correct -- Index...

## Next Steps

- Fix Dates

 - as.Date()
 
```{r fixdates}
# Fix the dates!
AQbeijing$NewDate = as.Date(AQbeijing$date)
str(AQbeijing)
head(AQbeijing)
```


```{r}
plot(pm25 ~ NewDate, data=AQbeijing)

```


 
- Select Data Before and After Lockdown
- Different for each city? Yes 
- 80 days

  - subset()
  
  
```{r subset}
# Dates for Beijing: "2020-02-10" Start

AQbeijingLockDown = as.Date("2020-02-10"); str(AQbeijingLockDown)

AQbeijingBefore = subset(AQbeijing, 
        subset=(NewDate > AQbeijingLockDown-80 &
                NewDate < AQbeijingLockDown))
AQbeijingAfter = subset(AQbeijing, 
        subset=(NewDate >= AQbeijingLockDown &
                NewDate < AQbeijingLockDown + 80))

plot(pm25~NewDate, AQbeijingBefore)
plot(pm25~NewDate, AQbeijingAfter)

```

- t-test Before and After
  - t.test()
  
  
```{r ttest}
#run ttest
t.test(AQbeijingBefore$pm25, AQbeijingAfter$pm25)


```

```{r ttest}
#extract mean_before, mean_after, p-value
ttest = t.test(AQbeijingBefore$pm25, AQbeijingAfter$pm25)
ttest$estimate
ttest$p.value


#export whole ttest
capture.output(ttest, file = "Welch_output.csv", append = TRUE)

#export just city, estimate, pvalue
estimate = ttest$estimate
pvalue = ttest$p.value
city = "Beijing"
capture.output(city, estimate, pvalue, file = "output.csv", append = TRUE)



#dump(city, estimate, pvalue, file = "output.csv", append=TRUE)

#save(city, estimate, pvalue, "output.csv")

#write.csv("Beijing", file="output.csv", row.names=FALSE, append=TRUE)


#write.csv(estimate, pvalue, file = "output.csv")

#save(estimate, pvalue, list = character(),
     #file = ("output.csv")

#save(estimate, pvalue, file = "output.RData")
#head("output.RData")
```

#repeating import data, scatterplot, fix dates, welch 2-sample t test, output for other cities.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Data

The data came from the Air Quality Historical Data Platform, an open data framework which gathers data from worldwide EPA monitering programs.

```{r cars}
#airqualitybeijing.csv <- #"/home/CAMPUS/mwl04747/github/China_air_pollution_COVID/beijing-air-quality.csv"


xianning.csv = "/home/CAMPUS/dapa2019/R/China_air_pollution_COVID/data/e/xianning-air-quality.csv"

AQ <- read.csv(xianning.csv)

str(AQ)

```

## Scatter Plot

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(AQ$pm25)
```

Dates aren't correct -- Index...

## Next Steps

- Fix Dates

 - as.Date()
 
```{r fixdates}
# Fix the dates
AQ$NewDate = as.Date(AQ$date)
str(AQ)
head(AQ)
```


```{r}
plot(pm25 ~ NewDate, data=AQ)

```


 
- Select Data Before and After Lockdown
- Different for each city? Yes 
- 80 days

  - subset()
  
  
```{r subset}
# Dates for city: "yyyy-mm-dd" Start Lockdown

AQLockDown = as.Date("2020-01-24"); str(AQLockDown)

AQxianningBefore = subset(AQ, 
        subset=(NewDate > AQLockDown-80 &
                NewDate < AQLockDown))
AQxianningAfter = subset(AQ, 
        subset=(NewDate >= AQLockDown &
                NewDate < AQLockDown + 80))

plot(pm25~NewDate, AQxianningBefore) 
plot(pm25~NewDate, AQxianningAfter)

```

- t-test Before and After
  - t.test()
  
  
```{r ttest}
#run ttest
t.test(AQxianningBefore$pm25, AQxianningAfter$pm25)


```

```{r ttest}
#extract mean_before, mean_after, p-value
ttest = t.test(AQxianningBefore$pm25, AQxianningAfter$pm25)
ttest$estimate
ttest$p.value


#export whole ttest
capture.output(ttest, file = "Welch_output.csv", append = TRUE)

#export just city, estimate, pvalue
estimate = ttest$estimate
pvalue = ttest$p.value
city = "Xianning"
capture.output(city, estimate, pvalue, file = "output.csv", append = TRUE)



#dump(city, estimate, pvalue, file = "output.csv", append=TRUE)

#save(city, estimate, pvalue, "output.csv")

#write.csv("Beijing", file="output.csv", row.names=FALSE, append=TRUE)


#write.csv(estimate, pvalue, file = "output.csv")

#save(estimate, pvalue, list = character(),
     #file = ("output.csv")

#save(estimate, pvalue, file = "output.RData")
#head("output.RData")
```








- BACI "Before and Ater Control-Impact"
  - aov() 
  
  
visualization:

- time series -- with abreak B/A
- box plot B/A





